---
title: "Lab 1: Exploratory Analysis of CEO Salary Data"
author: "Carmen Easterwood & Andrew Kabatznick"
date: "January 29, 2016"
output:
  pdf_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
attach(CEO)
```

# Introduction

Aaa

```{r}
```

# Univariate Analysis of Key Variables
### Years With Company
Summary Statistics for Years with Company
```{r}
summary(comten)
```

Variance of Years with Company
```{r}
var(comten)
```

Standard Deviation of Years with Company
```{r}
sd(comten)
```

Histogram of Years with Company
```{r}
hist(comten,main="Years with Company",xlab = "Years")
```

### Years as CEO
Summary Statistics for Years as CEO
```{r}
summary(ceoten)
```

Variance of Years as CEO
```{r}
var(ceoten)
```

Standard Deviation of Years as CEO
```{r}
sd(ceoten)
```

Histogram of Years as CEO
```{r}
hist(ceoten,main="Years as CEO",xlab = "Years")
```

### Profits
Summary Statistics for Profits
```{r}
summary(profits)
```

Variance of Profits
```{r}
var(profits)
```

Standard Deviation of Profits
```{r}
sd(profits)
```

Histogram of Profits
```{r}
hist(profits,main="1990 Profits",xlab = "millions",breaks = 100)
```

### Market Value
Summary Statistics for Market Value
```{r}
summary(mktval)
```

Variance of Market Value
```{r}
var(mktval)
```

Standard Deviation of Market Value
```{r}
sd(mktval)
```

Histogram of Market Value
```{r}
hist(mktval,main="Market Value at the End of 1990",xlab = "millions",breaks = 100)
```


# Key Bivariate Relationships

Aaa

### Relationship 1

Aaa
```{r}
```

# Possible Secondary Variables

Aaa
```{r}
```

# Potential Confounding Effects

Aaa
```{r}
```


# Variable Coding Issues and Missing Values

Aaa
```{r}
```

# Conclusion

Aaa
```{r}
```

