---
title: "Lab 1: Exploratory Analysis of CEO Salary Data"
author: "Carmen Easterwood & Andrew Kabatznick"
date: "January 29, 2016"
output:
  pdf_document: default
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

Aaa

```{r}
setwd("~/Desktop/MIDS/Statistics/stats_lab1")
ceosal <- load("ceo_w203.RData", ceo.env <- new.env())
ceo.df <- ceo.env[["CEO"]]
```

# Univariate Analysis of Key Variables
In the following section we will perform a univariate analysis of each of the variables in this dataset. Salary is our outcome variable, and profits and mktval are the key measures of market performance.

### Salary

CEO salary distribution is strongly skewed right.

```{r}
hist(ceo.df$salary, main = "Histogram of CEO Salary in 1990",
     xlab = "Salary in $ thousands", breaks = seq(0, 6000, by = 500))
axis(1, at = seq(0, 6000, by = 1000))
```

Median salary is $`r median(ceo.df$salary)` thousand, and there is one extreme outlier at $`r round(max(ceo.df$salary) / 1000, 1)` million.

```{r}
summary(ceo.df$salary)
```

### Age

CEO age peaks between 50 and 65 years old, but ranges all the way from `r min(ceo.df$age)` to `r max(ceo.df$age)`.
```{r}
hist(ceo.df$age, breaks = 14, main = "Histogram of CEO Age", xlab = "Age")
summary(ceo.df$age)
```

The variance is large:
```{r}
var(ceo.df$age)
```

### College

College is a dummy variable that takes a value of 1 if the CEO is a college graduate and 0 otherwise.

```{r}
pct.college <- (sum(ceo.df$college) / length(ceo.df$college))
```

`r round(pct.college * 100, 1)`% of the CEOs in this dataset are college graduates.

### Grad

Grad is a dummy variable that takes a value of 1 if the CEO holds an advanced degree and 0 otherwise.

```{r}
pct.grad <- (sum(ceo.df$grad) / length(ceo.df$grad))
```

`r round(pct.grad * 100, 1)`% of the CEOs in this dataset are college graduates.

```{r, include = FALSE}
sub <- subset(ceo.df, ceo.df$college == 0 & ceo.df$grad == 1)
```

**Note: Should we mention there are `r length(sub$salary)` CEOs with an advanced degree but no college degree? (Can that be correct?)**

### Years With Company
Summary Statistics for Years with Company
```{r}
summary(ceo.df$comten)
```

Variance of Years with Company
```{r}
var(ceo.df$comten)
```

Standard Deviation of Years with Company
```{r}
sd(ceo.df$comten)
```

Histogram of Years with Company
```{r}
hist(ceo.df$comten, main="Years with Company", xlab = "Years")
```

### Years as CEO
Summary Statistics for Years as CEO
```{r}
summary(ceo.df$ceoten)
```

Variance of Years as CEO
```{r}
var(ceo.df$ceoten)
```

Standard Deviation of Years as CEO
```{r}
sd(ceo.df$ceoten)
```

Histogram of Years as CEO
```{r}
hist(ceo.df$ceoten, main="Years as CEO", xlab = "Years")
```

### Profits
Summary Statistics for Profits
```{r}
summary(ceo.df$profits)
```

Variance of Profits
```{r}
var(ceo.df$profits)
```

Standard Deviation of Profits
```{r}
sd(ceo.df$profits)
```

Histogram of Profits
```{r}
hist(ceo.df$profits, main="1990 Profits",
     xlab = "millions", breaks = 100)
```

### Market Value
Summary Statistics for Market Value
```{r}
summary(ceo.df$mktval)
```

Variance of Market Value
```{r}
var(ceo.df$mktval)
```

Standard Deviation of Market Value
```{r}
sd(ceo.df$mktval)
```

Histogram of Market Value
```{r}
hist(ceo.df$mktval, main="Market Value at the End of 1990",
     xlab = "millions", breaks = 100)
```

# Key Bivariate Relationships

Aaa

### Relationship 1

Aaa
```{r}
```

# Possible Secondary Variables

Aaa
```{r}
```

# Potential Confounding Effects

Aaa
```{r}
```


# Variable Coding Issues and Missing Values

Aaa
```{r}
```

# Conclusion

Aaa
```{r}
```
