---
title: "Lab 1: Exploratory Analysis of CEO Salary Data"
author: "Carmen Easterwood & Andrew Kabatznick"
date: "January 29, 2016"
output:
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

Salary is our outcome variable, and profits and mktval are the key measures of market performance.

```{r}
setwd("~/Desktop/MIDS/Statistics/stats_lab1")
ceosal <- load("ceo_w203.RData", ceo.env <- new.env())
ceo.df <- ceo.env[["CEO"]]
```

# Univariate Analysis of Key Variables

In this section we will analyze each variable individually.

### Salary

CEO salary distribution is strongly skewed right.

```{r}
hist(ceo.df$salary, main = "Histogram of CEO Salary in 1990",
     xlab = "Salary in $ thousands", breaks = seq(0, 6000, by = 500))
axis(1, at = seq(0, 6000, by = 1000))
```

Median salary is $`r median(ceo.df$salary)` thousand, and there is one extreme outlier at $`r round(max(ceo.df$salary) / 1000, 1)` million.

```{r}
summary(ceo.df$salary)
```

### Age

CEO age peaks between 50 and 65 years old, but ranges all the way from `r min(ceo.df$age)` to `r max(ceo.df$age)`.
```{r}
hist(ceo.df$age, breaks = 14, main = "Histogram of CEO Age", xlab = "Age")
summary(ceo.df$age)
```

The variance of Age is large:
```{r}
var(ceo.df$age)
```

### College Degree

College is a dummy variable that takes a value of 1 if the CEO is a college graduate and 0 otherwise.

```{r}
pct.college <- (sum(ceo.df$college) / length(ceo.df$college))
```

`r round(pct.college * 100, 1)`% of the CEOs in this dataset are college graduates.

### Graduate Degree

Grad is a dummy variable that takes a value of 1 if the CEO holds an advanced degree and 0 otherwise.

```{r}
pct.grad <- (sum(ceo.df$grad) / length(ceo.df$grad))
```

`r round(pct.grad * 100, 1)`% of the CEOs in this dataset hold advanced degrees.

### Years With Company
The Median (`r round(median(ceo.df$comten), 1)`) and Mean (`r round(mean(ceo.df$comten), 1)`) both indicate that in general CEOs have been with their companies around 21 years. However, the data is highly variable with a min of `r round(min(ceo.df$comten), 1)` and a max of `r round(max(ceo.df$comten), 1)`.
```{r}
summary(ceo.df$comten)
```

Variance of Years with Company
```{r}
var(ceo.df$comten)
```

The standard deviation for years with the company is ~r round(sd(ceo.df$comten), 1), which is large but unsurprising given the variablity of the dataset.
```{r}
sd(ceo.df$comten)
```

Histogram of Years with Company
```{r}
hist(ceo.df$comten, main="Years with Company", xlab = "Years")
```

### Years as CEO
The years as CEO variable exhibits a large right skew with a median of around `r round(median(ceo.df$comten), 1)`. This is not entirely suprising given CEOs tend to be established in their careers and thus older, which may explain while things tail off. We must also wonder if there is a large turnover in the first few years of a CEOs tenure as most CEOs are between 0-5 years in the position.

Summary Statistics for Years as CEO
```{r}
summary(ceo.df$ceoten)
```

Variance of Years as CEO
```{r}
var(ceo.df$ceoten)
```

Standard Deviation of Years as CEO
```{r}
sd(ceo.df$ceoten)
```

Histogram of Years as CEO
```{r}
hist(ceo.df$ceoten, main="Years as CEO", xlab = "Years")
```

### Profits
Companies in the sample tend to make a profit as shown by the median profit of `r round(mean(ceo.df$comten), 1)`. The mean is significanlty higher than the median driven by a notable outlier of `r round(max(ceo.df$comten), 1)`. While profits can be either positibe or negative, we found `r sum(ceo.df$profits == -1)` datapoints that are equal to negtive 1. Given this same value is found for the same CEOs in the market value variable, we beleive this is a bad value.

Summary Statistics for Profits
```{r}
summary(ceo.df$profits)
```


Variance of Profits
```{r}
var(ceo.df$profits)
```

Standard Deviation of Profits
```{r}
sd(ceo.df$profits)
```

Histogram of Profits
```{r}
hist(ceo.df$profits, main="1990 Profits",
     xlab = "millions", breaks = 100)
```

### Market Value

Unsurprisingly, the Market Value for companies has a right skew. The Min `r round(min(ceo.df$mktval), 1)` looks to be in error as Market Values should range from 0 to infinity. We found `r sum(ceo.df$mktval == -1)` datapoints that are equal to negtive 1, which appears to be a bad value.

Summary Statistics for Market Value
```{r}
summary(ceo.df$mktval)
```

Variance of Market Value
```{r}
var(ceo.df$mktval)
```

Standard Deviation of Market Value
```{r}
sd(ceo.df$mktval)
```

Histogram of Market Value
```{r}
hist(ceo.df$mktval, main="Market Value at the End of 1990",
     xlab = "millions", breaks = 100)
```

# Key Bivariate Relationships

In this section we analyze some key bivariate relationships. Below is a scatterplot matrix of these key variables.

```{r}
library(car)
scatterplotMatrix(~ salary + profits + mktval, data = ceo.df)
```

### Profits & Market Value

These are our two measures of company performance, so we confirm they are positively correlated.

```{r}
cor(ceo.df$profits, ceo.df$mktval)
```

For the most part, as profits increase, so does market value.

```{r}
plot(ceo.df$profits, ceo.df$mktval,
     main = "Profits vs. Market Value in 1990",
     xlab = "Market Value ($ millions)",
     ylab = "Profits ($ millions)")
abline(lm(ceo.df$mktval ~ ceo.df$profits), col = "red")
```

### Profits & CEO Salary
Salary and Profits don't have a strong linear relationship as demonstrated by the low correlation.
```{r}
cor(ceo.df$salary, ceo.df$profits)
```


When profits are between -500 million dollars and 500 million dollars and salary is less than 2 million dollars, there seems to be a slight linear relationship. However, outlier CEOs, who significantly outperformed their peers by making significant profits per dollar earned, skew the linearity of the data.
```{r}
plot(ceo.df$salary, ceo.df$profits,
     main = "Salary vs. Profits in 1990",
     xlab = "Salary ($ thousands)",
     ylab = "Profits ($ millions)")
```

However, after performing a logorithmic transformation on the profits variable, the variables show more a linear relationship.   
```{r}
plot(ceo.df$salary, log(ceo.df$profits),
     main = "Salary vs. Profits in 1990",
     xlab = "Salary ($ thousands)",
     ylab = "Log of Profits ($ millions)")
```

### Market Value & CEO Salary
Like the relationship between salary and profits, salary and market value also don't have a strong linear relationship, which can be seen in the correlation below. Given the high correlation between profits and market value, it is thus unsurprising that salary and market value are also not highly correlated.
```{r}
cor(ceo.df$salary, ceo.df$mktval)
```

Similar to salary and profits, there appears to be some linearity when profits and market value are low, but as market value rises the linear reltionship weakens.
```{r}
plot(ceo.df$salary, ceo.df$mktval,
     main = "Salary vs. Market Value in 1990",
     xlab = "Salary ($ thousands)",
     ylab = "Market Value ($ millions)")
```

Similar to the profit variable, a logorthmic transformation on the market value variable demonstrates a more linear relationship between the two variables.
```{r}
plot(ceo.df$salary, log(ceo.df$mktval),
     main = "Salary vs. Market Value in 1990",
     xlab = "Salary ($ thousands)",
     ylab = "Market Value ($ millions)")
```
# Possible Secondary Variables
Some secondary variables that may affect salary are the CEO's age, tenure, and education level. Below is a scatterplot of salary on age and tenure variables. Since education level is split into two dummy variables, we do not find a scatterplot of it to be useful.

```{r}
scatterplotMatrix(~ salary + age + comten + ceoten, data = ceo.df)
```

### Salary vs. Age
Age has a slight positive relationship with CEO salary, likely because older CEOs have more work experience. However, with a correlation of only `r cor(ceo.df$salary, ceo.df$age)`, age is not a key variable in this analysis.

```{r}
plot(ceo.df$age, ceo.df$salary,
     main = "Salary vs. Age",
     xlab = "Age",
     ylab = "Salary ($ thousands)")
abline(lm(ceo.df$salary ~ ceo.df$age), col = "red")
```

### Salary vs. Education
We assign each CEO to one of three education levels: Advanced Degree, College Graduate, or Less than College. There are 2 cases of CEOs with an advanced degree but no college degree, and these are assigned to the "Advanced Degree" level.

```{r}
educLevelFunc <- function(college, grad) {

    if (grad == 1) {retStr = "Advanced Degree"}
    else if (college == 1) {retStr = "College Graduate"}
    else {retStr = "Less than College"}

    return(retStr)
}

ceo.df$educLevel <- mapply(educLevelFunc, ceo.df$college, ceo.df$grad)
table(ceo.df$educLevel)
```

Following is a boxplot of CEO salary by education level. The distribution category for the "less than college" category is not reliable since it only has 5 data points. However, we can see that CEOs with an advanced degree have approximately the same salary distribution as CEOs with only a college degree.

```{r}
boxplot(salary ~ educLevel, data=ceo.df,
     main = "Boxplot of Salary by Education Level",
     ylab = "Salary ($ thousands)")
```

### Salary vs. Tenure

We have two tenure variables: CEO tenure and company tenure. These variables are closely related, since company tenure must rise whenever CEO tenure rises. Note that for CEOs who are hired from outside the company, CEO tenure and company tenure are the same.

```{r}
outside.ceo <- subset(ceo.df, ceo.df$ceoten == ceo.df$comten)
pct.outside.ceo <- (length(outside.ceo$ceoten) / length(ceo.df$ceoten))
```

(*Note:* In this dataset, `r round(pct.outside.ceo * 100, 1)`% of CEOs are brought in from outside the company.)  

As shown in the scatterplots below, CEO salary has essentially no relationship with the number of years the CEO has been with the company, and only a slight positive relationship with the number of years as CEO.

```{r}
cor(ceo.df$salary, ceo.df$comten)
plot(ceo.df$comten, ceo.df$salary,
     main = "Salary vs. Years with the Company",
     xlab = "Years with the Company",
     ylab = "Salary ($ thousands)")
abline(lm(ceo.df$salary ~ ceo.df$comten), col = "red")
cor(ceo.df$salary, ceo.df$ceoten)
plot(ceo.df$ceoten, ceo.df$salary,
     main = "Salary vs. Years as CEO",
     xlab = "Years as CEO",
     ylab = "Salary ($ thousands)")
abline(lm(ceo.df$salary ~ ceo.df$ceoten), col = "red")
```


# Potential Confounding Effects

Aaa
```{r}
```


# Variable Coding Issues and Missing Values

*Mention 2 CEOs with advanced degree but not college degree*  
*Mention 1 CEO with CEO tenure > company tenure*  
*Mention 5 Companies with market cap equal to -1*  
```{r}
```

# Conclusion

Aaa
```{r}
```
